---
title: "Recapture Analsyis: COLONY"
author: "John D. Robinson"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
linkcolor: blue
---

# Recapture Analysis in `COLONY`
The `COLONY` job ran successfully in ~10 hours (using the parallel `COLONY` executable with 8 cores). Note that this analysis is faster when using the monogamous mating model, which [Wang (2016)](https://doi.org/10.1111/1755-0998.12452) recommends. This way, only full-sib groups are identified at first. It is unlikely that pairs of samples identified as half-sibs would actually be recaptures. While this is fine for the clone analysis in `COLONY`, it would not be appropriate for an analysis intended to reconstruct the pedigree of the sample or estimate $N_b$ using the sibship method implemented in `COLONY`. Those analyses would require the polygamous mating model.  

In this markdown, we read in the outputs from `COLONY` and calculate a few basic summaries...  
```{r readCOLout}
col.out <- read.table("/home/jdrobins/CGA_BlackBears/recaps/CGA.BestClone", 
                      header = TRUE)

col.list <- vector("list", nrow(col.out))
for(i in 1:nrow(col.out)) {
  col.list[[i]] <- strsplit(col.out[i,3], split=",")[[1]]
}

col.list[[1]]
col.list[[2]]
col.list[[3]]
```

## Summarize output
Some quick summaries of the clusters identified by `COLONY`...  
```{r summCOLONY}
#Total number of unique genotypes
length(col.list)

#Histogram of cluster probabilities
hist(col.out$Prob, xlab = "Cluster Probability")

#Sizes of clusters
nsamps <- sapply(col.list, FUN=length)
hist(nsamps, xlab = "Number of Samples per Cluster", breaks = 20)

#Relationship between cluster size and probability?
plot(x = nsamps, y=col.out$Prob, 
     xlab = "Cluster size", ylab = "Cluster probability",
     pch = 19, cex = 0.5)
cor.test(nsamps, col.out$Prob)
```

Further analysis, including assignment of cloneIDs and comparisons of outputs from `COLONY` and `PLINK2` are in `10_recapFinal`.  

---


