---
title: "Create Capture History Objects"
author: "John D. Robinson"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
linkcolor: blue
---

# Building Capture Histories for SCR Analyses
In this markdown, we'll create capture histories for the SCR analyses that follow. Each object will include i) a `y` array of capture histories (with `dim = c(i, j, k, t)`, individuals, traps, occasions, years), ii) a `pc` array (of similar dimensions) with previous capture information (for models with a local behavioral effect), and iii) a `clone` vector that has identifiers for each individual $i$ (from the `finalClone` column of our `sampleMetadata` file). We want to build several of these objects, given interest in fitting different (open vs. closed) models. 

1. For closed population models - data for two years saved separately (3-d array), separate for the two sexes as well
    + Females in 2023
    + Females in 2024
    + Males in 2023
    + Males in 2024
2. For open population models - two years combined (4-d array), but still modeling the sexes separately
    + Females (2023 and 2024)
    + Males (2023 and 2024)

## Load data, source function
Load *i*) the sample metadata with clone ID information and *ii*) the `.Rda` with locations and yearly operational status of hair snares. Source in functions for *i*) building capture history and previous capture objects and *ii*) calculating summaries of capture histories from a 3-d or 4-d `y` array (based on a function from the [`scrbook` R package](https://github.com/jaroyle/scrbook)).  
```{r loadData}
#Load sample metadata (smd)
smd <- read.csv("../sampleinfo/sampleMetaData_cloneID.csv", header = TRUE)
#Load snare location information
load("../sampleinfo/snareLocations.Rda")
ls()

#Source the makeCH function
source("../input/makeCH.R")
#Source the sumstats functions (for 3d and 4d arrays)
source("../input/sumstats_4d.R")
```

## Closed population models
Splitting the dataset up by year and sex...  
```{r makeCHclosed}
#Start with all hair samples from 2023 and 2024
sampsALL <- smd[which(!is.na(smd$finalClone) &
                        smd$is.SCR == TRUE & 
                        smd$year %in% c(2023,2024) &
                        smd$FILT == FALSE &
                        !is.na(smd$finalSEX)),]

#Check the dataset
nrow(sampsALL)
table(sampsALL$region)
table(sampsALL$finalSEX)
sum(!sampsALL$finalSEX %in% c("M","F"))
length(unique(sampsALL$finalClone[!is.na(sampsALL$finalClone)]))
sum(is.na(sampsALL$finalClone))

#Separately for males, females, 2023, and 2024
snares23 <- locs[oper[,1] == 1,]
snares24 <- locs[oper[,2] == 1,]

#M23
Msamps23 <- sampsALL[which(sampsALL$year == 2023 &
                       sampsALL$finalSEX == "M"),]
dim(Msamps23)
table(Msamps23$finalSEX)
table(Msamps23$year)
length(unique(Msamps23$finalClone[!is.na(Msamps23$finalClone)]))

M23 <- makeCH(samps=Msamps23,
              traps=snares23,
              yrs=2023)
dim(M23$y)
dim(M23$pc)
length(M23$clone)

#M24
Msamps24 <- sampsALL[which(sampsALL$year == 2024 &
                       sampsALL$finalSEX == "M"),]
dim(Msamps24)
length(unique(Msamps24$cloneID[!is.na(Msamps24$cloneID)]))

M24 <- makeCH(samps=Msamps24,
              traps=snares24,
              yrs=2024)
dim(M24$y)
dim(M24$pc)
length(M24$clone)

length(unique(c(M23$clone,M24$clone)))

#F23
Fsamps23 <- sampsALL[which(sampsALL$year == 2023 &
                       sampsALL$finalSEX == "F"),]
dim(Fsamps23)
length(unique(Fsamps23$finalClone[!is.na(Fsamps23$finalClone)]))


F23 <- makeCH(samps=Fsamps23,
              traps=snares23,
              yrs=2023)
dim(F23$y)
dim(F23$pc)
length(F23$clone)

#F24
Fsamps24 <- sampsALL[which(sampsALL$year == 2024 &
                       sampsALL$finalSEX == "F"),]
dim(Fsamps24)
length(unique(Fsamps24$finalClone[!is.na(Fsamps24$finalClone)]))

F24 <- makeCH(samps=Fsamps24,
              traps=snares24,
              yrs=2024)
dim(F24$y)
dim(F24$pc)
length(F24$clone)

length(unique(c(F23$clone, F24$clone)))

#And the summary stats
rowSums(M23$y)
sumstats_3d(M23$y)
rowSums(M24$y)
sumstats_3d(M24$y)

rowSums(F23$y)
sumstats_3d(F23$y)
rowSums(F24$y)
sumstats_3d(F24$y)
```

## Open population models
Combining data for years, but sexes are still separate...  
```{r makeCHopen}
#Females - both years
Fsamps <- sampsALL[which(sampsALL$finalSEX == "F"),]
length(unique(Fsamps$finalClone[!is.na(Fsamps$finalClone)]))

Fch <- makeCH(samps=Fsamps,
              traps=locs,
              yrs=c(2023,2024))
dim(Fch$y)
dim(Fch$pc)
length(Fch$clone)

#Males - both years
Msamps <- sampsALL[which(sampsALL$finalSEX == "M"),]
length(unique(Msamps$finalClone[!is.na(Msamps$finalClone)]))

Mch <- makeCH(samps=Msamps,
              traps=locs,
              yrs=c(2023,2024))

dim(Mch$y)
dim(Mch$pc)
length(Mch$clone)

#And the summary stats
rowSums(Fch$y)
sumstats_4d(Fch$y)

rowSums(Mch$y)
sumstats_4d(Mch$y)
```

## Save the output
Save all of these objects (`F23`,`F24`,`M23`,`M24`,`Fch`,`Mch`) to a single `.Rda` file...  
```{r saveOut}
save(F23, F24, M23, M24, Fch, Mch, file = "CGA_CH.Rda")
```

---

