---
title: "Fitting the Hooker et al. (2020) model to both sexes"
author: "John D. Robinson"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
linkcolor: blue
---

# Fitting a model similar to that from Hooker et al. (2020) with *both sexes*
The manuscript by [Hooker et al. (2020)](https://doi.org/10.1002/jwmg.21887) fit an open-population SCR model to recapture data from female black bears in central GA using hair samples collected over a period of 5 years (2012-2016). They compared several models (see their Table 2), but the best-fit model included a constant survival rate across years, density-dependent per capita recruitment, year-specific baseline detection probabilities, and a local behavioral response to capture. Estimates of abundance from their analysis ranged from 106 to 166 across the study period.  

Here, we attempt to use Hooker et al. (2020)'s code (from [github.com/rbchan/ga-bear-pva](https://github.com/rbchan/ga-bear-pva)) to fit a very similar model to data hair samples collected in 2023 and 2024 (from *both male and female bears*) in the same area. In particular, the analysis below draws heavily from their [fit.R](https://github.com/rbchan/ga-bear-pva/blob/master/R/fit.R) script and their [`JAGS` model](https://github.com/rbchan/ga-bear-pva/blob/master/R/dT-s0-rPDD2-phi0-pYB-sig0.jag), and uses their [state space `.tif` file](https://raw.githubusercontent.com/rbchan/ga-bear-pva/refs/heads/master/data/state-space360.tif).  

The`JAGS` model from the Hooker manuscript is modified extensively below. Specifically, we remove the density-dependence of per-capita recruitment (with only 1 recruitment period, this should have little effect on the model). We also fit the model to data from both sexes and include several sex-specific parameters. We model differences in the sexes for the following: 

i. $p_0$ - baseline detection probability, probability of detection when an individual's activity center is in the same location as a hair snare
ii. `behave` - the (local) behavioral effect of capture, assumed to increase $p_0$ for individuals that have previously been detected in a specific location 
iii. $\sigma$ - the decay rate of detection probability with distance (related to movement)
iv. $\phi$ - (apparent) survival rate between primary sampling periods (years)
v. $\psi$ - the data augmentation parameter

We augment the male and female capture histories separately, so much of the code in the model is duplicated. This also means we have separate objects in the model for the male and female activity centers, individual alive/dead/available statuses, and so on. We will need to initialize these variables separately for males and females, so there will be a bit more to organize in the `JAGS data` and `inits` portions.  

Here's the `JAGS` model we're using...  
```{r JAGSmodelCode}
writeLines(readLines("../input/HookerModel-noDensDep_bothSexes2.jag"))
```

## Load the data
We do NOT need to create a combined capture history object, since we augment male and female capture histories separately in our model.  
```{r loadData}
#Load and collect the necessary data objects
load("CGA_CH.Rda")
load("../sampleinfo/snareLocations.Rda")
ls()

#Some summary stats
source("../input/sumstats_4d.R")

#capture history lists have 3 elements:
# y - the capture history, 0 for no detection, 1 for detection
# pc - previous capture, has individual i been seen in trap j before?
# clone - clone indices (finalClone) that correspond to rows of y and pc
names(Mch)
names(Fch)

#Check that j, k, and t dimensions of the y and pc arrays in Mch and Fch are =
dim(Mch$y)
dim(Fch$y)
dim(Mch$pc)
dim(Fch$pc)

sumstats_4d(Fch$y)
sumstats_4d(Mch$y)
```

## Rerun `fit.R`
From here on out, we'll follow along with the [fit.R](https://github.com/rbchan/ga-bear-pva/blob/master/R/fit.R) script from Hooker et al. (2020). Note that there will be a number of modifications from fit.R in the following chunks. There are several intermediate sample collections in that script that are not implemented here (for WAIC calculations, forecasting, etc.). Note also that most of the markdown is shared with the other SCR analyses (*e.g.*, `14_fitHooker`, `15_fitHookerM`, `16_fitHooker_move`, etc.).  
```{r fitRscript}
## Fit open population SCR models to central GA black bear data
## Female-only analysis

library(rjags)
library(lattice)
library(raster)

# For consistency, rename the objects to match their scheme... 
ch4d.f <- Fch$y
prevcap.f <- Fch$pc
ch4d.m <- Mch$y
prevcap.m <- Mch$pc
traps <- locs


## Inspect the loaded data objects
str(ch4d.f)
str(ch4d.m)
str(traps)

## Check that traps are in the same order in both objects
## Must be TRUE
all(rownames(traps)==dimnames(ch4d.f)[[3]])
all(rownames(traps)==dimnames(ch4d.m)[[3]])
all(rownames(oper)==rownames(traps))


## Format data

yF <- ch4d.f ## 4D encounter histories
str(yF)
yM <- ch4d.m
str(yM)

y.dim <- dim(yF)
nFem <- dim(yF)[1]
nMale <- dim(yM)[1]
nTraps <- dim(yF)[2]
nWeeks <- dim(yF)[3]
nYears <- dim(yF)[4]



## Females
Mf <- 500 # Data aug. Maximum # offemales that could have been alive
all(rownames(prevcap.f)==rownames(yF)) ## Must be TRUE

## Non-spatial, robust design format
yF.nonsp.robust <- apply(yF>0, c(1,3,4), any)*1L
str(yF.nonsp.robust)

## Non-spatial, non-robust format
yF.nonsp.nonrobust <- apply(yF.nonsp.robust>0,
                           c(1,3), any)*1L
str(yF.nonsp.nonrobust)

## Spatial, non-robust format
yF.sp.nonrobust <- apply(yF>0, c(1,2,4), any)*1L
str(yF.sp.nonrobust)

## First and last detections 
first.detF <- apply(yF.nonsp.nonrobust>0, 1, function(x) min(which(x)))
last.detF <- apply(yF.nonsp.nonrobust>0, 1, function(x) max(which(x)))
known.aliveF <- last.detF-first.detF+1
table(known.aliveF)

## Data augmentation
yFaug <- array(0, c(Mf, nTraps, nWeeks, nYears))
yFaug[1:nFem,,,] <- yF

str(yFaug)


## Known values of the alive/dead state matrix
zFdata.aug <- matrix(NA, Mf, nYears)
for(i in 1:nFem) {
    zFdata.aug[i,first.detF[i]:last.detF[i]] <- 1
}

## Visualize the observed information about lifetime
image(t(zFdata.aug[1:nFem,]))

## Bears known to be alive in each year
colSums(zFdata.aug[1:nFem,], na.rm=TRUE)



## Males
Mm <- 500 # Data aug. Maximum # of males that could have been alive
all(rownames(prevcap.m)==rownames(yM)) ## Must be TRUE

## Non-spatial, robust design format
yM.nonsp.robust <- apply(yM>0, c(1,3,4), any)*1L
str(yM.nonsp.robust)

## Non-spatial, non-robust format
yM.nonsp.nonrobust <- apply(yM.nonsp.robust>0,
                           c(1,3), any)*1L
str(yM.nonsp.nonrobust)

## Spatial, non-robust format
yM.sp.nonrobust <- apply(yM>0, c(1,2,4), any)*1L
str(yM.sp.nonrobust)

## First and last detections 
first.detM <- apply(yM.nonsp.nonrobust>0, 1, function(x) min(which(x)))
last.detM <- apply(yM.nonsp.nonrobust>0, 1, function(x) max(which(x)))
known.aliveM <- last.detM-first.detM+1
table(known.aliveM)

## Data augmentation
yMaug <- array(0, c(Mm, nTraps, nWeeks, nYears))
yMaug[1:nMale,,,] <- yM

str(yMaug)


## Known values of the alive/dead state matrix
zMdata.aug <- matrix(NA, Mm, nYears)
for(i in 1:nMale) {
    zMdata.aug[i,first.detM[i]:last.detM[i]] <- 1
}

## Visualize the observed information about lifetime
image(t(zMdata.aug[1:nMale,]))

## Bears known to be alive in each year
colSums(zMdata.aug[1:nMale,], na.rm=TRUE)



## Operational status of traps
nTrapsYr <- colSums(oper)
nTrapsYr
oper.traps <- matrix(0L, nTraps, nYears)
for(i in 1:nYears) {
    oper.traps[1:nTrapsYr[i],i] <- which(oper[,i]==1)
}


## State-space
x.range <- range(traps[,1])
y.range <- range(traps[,2])
buffer <- 4000
xlim0 <- x.range+c(-buffer, buffer)
ylim0 <- y.range+c(-buffer, buffer)

#!# Need to point to the location where I've saved this file
S360 <- raster("../input/state-space360.tif")

#Convert to UTM-17N-WGS84 to match snareLocations.Rda
newproj <- "+proj=utm +zone=17 +datum=WGS84 +units=m +no_defs +type=crs"
S360 <- projectRaster(S360, crs=newproj)

traps.in <- round(extract(S360, traps))

plot(S360)

## Convert raster to a data.frame
S360dat <- as.data.frame(S360, xy=TRUE)
names(S360dat) <- c("x", "y", "S360")
S360dat$x <- floor(S360dat$x)
S360dat$y <- floor(S360dat$y)
table(diff(S360dat$x))
table(diff(S360dat$y))

nPix360 <- nrow(S360dat)
dist360 <- matrix(NA, nPix360, nTraps)
for(j in 1:nTraps) {
    dist360[,j] <- sqrt((S360dat$x-traps[j,1])^2 +
                        (S360dat$y-traps[j,2])^2)
}

S360dat$pixIn <- (rowSums(dist360<buffer)>0)*1L
table(S360dat$pixIn)

dim.S360 <- dim(S360)

S360dat$habitat <- ifelse((S360dat$S360>0) &
                          (S360dat$pixIn>0),
                          1L, 0L)
```

Before proceeding further, it's important to note that the trap locations used in this study were not identical to those from Hooker et al. (2020). So it won't be surprising if we run into some issues where our traps are located outside of their state space. With that said, let's take a look at how our traps align.  

Sure enough, there were some slight disagreements here. The `S` object ended up with some `NA` entries that led to errors in the **JAGS** run. Replacing those with `0` or `1` seems to get around the issue. It seems that these are cells that were not considered in the Hooker et al. (2020) study, so we don't have any information on habitat characteristics. With that in mind, it's more conservative to assume that they are NOT habitat, than to allow bear activity centers to be located in these cells.  
```{r fitscript_cont1, crop=NULL}
## Visualize traps and state space
levelplot(habitat ~ x+y, S360dat, aspect='iso',
          panel=function(...) {
              panel.levelplot(...)
              lpoints(traps, pch=3)
          })


delta <- res(S360)[1]
xseq <- sort(unique(S360dat$x))
yseq <- sort(unique(S360dat$y), decreasing=TRUE)
xlim <- range(xseq)+c(-delta/2,delta/2)
ylim <- range(yseq)+c(-delta/2,delta/2)

dim.S360
range(ceiling((runif(1e5, xlim[1], xlim[2])-xlim[1])/delta))
range(ceiling((ylim[2]-runif(1e5, ylim[1], ylim[2]))/delta))


S <- matrix(S360dat$habitat, dim.S360[1], dim.S360[2],
            dimnames=list(yseq, xseq), byrow=TRUE)
str(S)

plot(raster(S))

#Area of the state space
sum(S, na.rm=TRUE)*delta^2 / 1e6
Sarea <- sum(S, na.rm=TRUE)*delta^2 / 1e6

#!# Where are these NA values??
Stmp <- S
Stmp[is.na(Stmp)] <- 500
plot(raster(Stmp))

#!# These appear to be points outside of the polygon assessed for habitat
#!# In the Hooker et al. (2020) study
#!# There are relatively few of these, and they are on the periphery of S
#!# Better to assume they're NOT habitat than to assume that they are

#!# Need to recode them to 0's or 1's for this to run #!#
S[is.na(S)] <- 0
```

The `NA` entries that caused problems for **JAGS** seem to be only along the periphery of the state space. Seems likely that these crop up because of slightly different positions of the traps which meant that the 4km buffer would've picked up a few additional cells in the raster.  

## Fit the model
The analysis in **JAGS** is below. First we prepare the data and initial values.  

```{r JAGSprep, crop=NULL}
## JAGS

library(rjags)



### Put the data in a list
jd <- list(yF=yF, yF.zero=matrix(0, Mf, nYears),
           yM=yM, yM.zero=matrix(0, Mm, nYears),
           zF=zFdata.aug,
           zM=zMdata.aug,
           nFem=nFem, nMale=nMale,
           Mf=Mf, Mm=Mm,
           nTrapsYr=as.integer(nTrapsYr),
           nWeeks=nWeeks, nYears=nYears, 
           oper.traps=oper.traps,
           prevcapF=prevcap.f+1,
           prevcapM=prevcap.m+1,
           S=S, delta=delta,
           onesF=rep(1, Mf),
           onesM=rep(1, Mm),
           x=traps, xlim=xlim, ylim=ylim,
           poly.area=Sarea)

str(jd)



#### Function to generate initial values
ji <- function() {
  nFem <- jd$nFem
  nMale <- jd$nMale
  Mf <- jd$Mf
  Mm <- jd$Mm
  nYears <- jd$nYears
  nWeeks <- jd$nWeeks
  traps <- jd$x
  nTraps <- nrow(traps)
  
  #AC and z for Females
  zFi <- matrix(0, Mf, nYears)
  sFi1 <- traps[sample.int(nTraps, size=Mf, replace=TRUE,
                          prob=traps.in),]
  sFi <- as.matrix(sFi1)
  for(i in 1:nFem) {
    zFi[i,first.detF[i]:last.detF[i]] <- NA
    traplocs.i <- (rowSums(jd$yF[i,,,])>0) * traps.in
    if(all(!traplocs.i))
      next
    sFi[i,] <- colMeans(traps[which(traplocs.i == 1),,drop=FALSE])
    #Check that the average detection location is in habitat
    testAC <- 1 == jd$S[round((jd$ylim[2]-sFi[i,2])/jd$delta+0.5),
                        round((sFi[i,1]-jd$xlim[1])/jd$delta+0.5)]
    S360hab <- S360dat[which(S360dat$habitat == 1),c("x","y")]
    if(!testAC) {
      tmp.dist <- sqrt((S360hab$x-sFi[i,1])^2 + 
                         (S360hab$y-sFi[i,2])^2)
      sFi[i,1] <- S360hab$x[which(tmp.dist == min(tmp.dist))]
      sFi[i,2] <- S360hab$y[which(tmp.dist == min(tmp.dist))]
      rm(tmp.dist)
    }
    rm(testAC)
  }
  
  #AC and z for males
  zMi <- matrix(0, Mm, nYears)
  sMi1 <- traps[sample.int(nTraps, size=Mm, replace=TRUE,
                           prob=traps.in),]
  sMi <- as.matrix(sMi1)
  for(i in 1:nMale) {
    zMi[i,first.detM[i]:last.detM[i]] <- NA
    traplocs.i <- (rowSums(jd$yM[i,,,])>0) * traps.in
    if(all(!traplocs.i))
      next
    sMi[i,] <- colMeans(traps[which(traplocs.i == 1),,drop=FALSE])
    #Check that the average detection location is in habitat
    testAC <- 1 == jd$S[round((jd$ylim[2]-sMi[i,2])/jd$delta+0.5),
                        round((sMi[i,1]-jd$xlim[1])/jd$delta+0.5)]
    S360hab <- S360dat[which(S360dat$habitat == 1),c("x","y")]
    if(!testAC) {
      tmp.dist <- sqrt((S360hab$x-sMi[i,1])^2 + 
                         (S360hab$y-sMi[i,2])^2)
      sMi[i,1] <- S360hab$x[which(tmp.dist == min(tmp.dist))]
      sMi[i,2] <- S360hab$y[which(tmp.dist == min(tmp.dist))]
      rm(tmp.dist)
    }
    rm(testAC)
  }

  p0i <- array(NA, c(2, nYears, 2))
  p0i[,,2] <- matrix(runif(2*nYears), 2, nYears)
  out <- list(zF=zFi, zM=zMi, sF=sFi, sM=sMi, p0st=p0i, 
              sigma=runif(2, 3000, 4000), 
              behave=runif(2, 0.5, 1),
              psi=matrix(c(runif(2), rep(NA, nYears)), 2, nYears),
              phi=runif(2), #!# Added this, we didn't draw init. phi before?
              gamma=runif(1, 0.2, 0.3)) 
  out$".RNG.name" <- "base::Mersenne-Twister" 
  out$".RNG.seed" <- sample.int(1e6, 1)
  return(out)
}

str(ji())

## Visualize initial values of activity centers
plot(raster(S, xmn=xlim[1], ymn=ylim[1], xmx=xlim[2], ymx=ylim[2]))
points(ji()$sF, pch = 1, cex = 0.5, col = "pink")
points(ji()$sM, pch= 1, cex = 0.5, col = "lightblue")
rect(xlim[1], ylim[1], xlim[2], ylim[2], border="blue")


##Focus in on a few individuals with multiple detections
# is ji() placing initial activity centers in the right places?
par(mar = c(3,3,3,3))
closer.look <- which(rowSums(jd$yF) > 5)
for(i in closer.look) {
  tmp.s <- ji()$sF[i,]
  plot(raster(S, xmn=xlim[1], ymn=ylim[1], xmx=xlim[2], ymx=ylim[2]), 
       main = paste0 ("Female ",i, " - ", "Clone ", Fch$clone[i]), 
       legend =FALSE)
  mtext(text = paste("Total Detections =", sum(jd$yF[i,,,])))
  points(traps, pch = 3, col = "red", cex = 0.5)
  points(jd$x[which(rowSums(jd$yF[i,,,]) > 0),], pch = 8, col = "blue", cex = 1.2)
  points(x=tmp.s[1], y=tmp.s[2], pch = 17, col = "grey")
  legend("topleft", bty = "n", pch = c(15, 3, 8, 17), cex = 0.75,
         col = c(terrain.colors(10)[1],"red","blue","grey"),
         legend=c("state space", "snares","detections","initial AC"))
  rm(tmp.s)
}


#### Parameters to monitor
jp <- c("p0st", "sigma", "behave", "gamma", "phi", "psi",
        "EN", "N", "ENf", "Nf", "ENm", "Nm", "Ntotal", "totF", "totM",
        "ED", "RD", "EDf", "RDf", "EDm", "RDm",
        "survivingF", "survivingM", "recruitF", "recruitM",
        "lambdaF", "lambdaM", "lambda")

```

Now to run the MCMC sampler. Note that several steps from the full [fit.R](https://github.com/rbchan/ga-bear-pva/blob/master/R/fit.R) script from [Hooker et al. (2020)](https://doi.org/10.1002/jwmg.21887) were skipped below. These steps involved calculation of WAIC or recording information needed for forecasting future population sizes under the model. Here, we simplify to two steps:

i. Adaptive phase sampling - 1000 iterations per chain 
ii. Collect samples from the posterior - 10,000 iterations per chain

```{r JAGSrun, crop=NULL}
## Models in parallel

library(parallel)

(nCores <- max(1, min(detectCores()-1, 4)))

cl1 <- makeCluster(nCores)

clusterExport(cl1, c("jd", "ji", "jp",
                     "traps.in", "S360dat",
                     "first.detF", "last.detF", "first.detM", "last.detM"))

## Compile and adapt

system.time({
    adapt <- clusterEvalQ(cl1, {
        library(rjags)
        load.module("dic")
        load.module("glm")
        mod <- jags.model(file="../input/HookerModel-noDensDep_bothSexes2.jag",
                         data=jd, inits=ji,
                         n.chains=1, n.adapt=1000) 
        return(mod)
    })
}) 

save(adapt, file="adapt-bothSexes2.gzip")




## Draw posterior samples, including the latent variables needed
## for forecasting
system.time({
  samples <- clusterEvalQ(cl1, {
    mod1 <- coda.samples(model=mod,
                         variable.names=c(jp, "deviance", 
                                          "zF", "aF", "sF", 
                                          "zM", "aM", "sM"),
                         n.iter=10000, n.thin=1) 
    return(as.mcmc(mod1))
  })
})



save(samples, file="samples-bothSexes2.gzip")

```

## Posterior samples

Now to plot a couple of posteriors and save the workspace for further analysis.  

```{r wrapUp, crop=NULL}
# Use the jc4.out object for posteriors
mc1 <- as.mcmc.list(samples)

focal.parms <- c("Nf[1]","Nf[2]","Nm[1]","Nm[2]",
                 "sigma[1]","sigma[2]","phi[1]","phi[2]","gamma",
                 "lambdaF", "lambdaM", "lambda",
                 "p0st[1,1,1]","p0st[1,1,2]","p0st[1,2,1]","p0st[1,2,2]",
                 "p0st[2,1,1]","p0st[2,1,2]","p0st[2,2,1]","p0st[2,2,2]",
                 "behave[1]","behave[2]")

# Posterior summary
summary(mc1[,focal.parms])

# Check ESS
effectiveSize(mc1[,focal.parms])

# Check Rhat
gelman.diag(mc1[,focal.parms])
```

Now plot the posterior distributions for the model parameters
```{r postPlot, crop=NULL}
# Plot some posteriors
for(p in 1:length(focal.parms)) {
  plot(mc1[,focal.parms[p]], main = focal.parms[p])
}
```

Now plot the density of inferred activity centers in the state space...  
```{r DensHooker, crop=NULL}
# Finally the density of activity centers in the study region
library(scrbook)
library(raster)
library(sf)
library(usmap)
library(prettymapr)

# Slightly modified plotting function
# See scrbook::SCRdensity() for the original
mySCRdensity <- function (obj, nx = 30, ny = 30, Xl = NULL, Xu = NULL, Yl = NULL, 
    Yu = NULL, scalein = 100, scaleout = 100, col = "gray", ncolors = 10, 
    whichguy = NULL) {
    Sxout <- obj$Sx
    Syout <- obj$Sy
    z <- obj$z
    niter <- nrow(z)
    if (is.null(Xl)) {
        Xl <- min(Sxout) * 0.999
        Xu <- max(Sxout) * 1.001
        Yl <- min(Syout) * 0.999
        Yu <- max(Syout) * 1.001
    }
    xg <- seq(Xl, Xu, , nx)
    yg <- seq(Yl, Yu, , ny)
    guy <- col(Sxout)
    Sxout <- cut(Sxout[z == 1], breaks = xg)
    Syout <- cut(Syout[z == 1], breaks = yg)
    if (is.null(whichguy)) {
        Dn <- table(Sxout, Syout)/niter
        area <- (yg[2] - yg[1]) * (xg[2] - xg[1]) * scalein
        Dn <- (Dn/area) * scaleout
    }
    else {
        Dn <- table(Sxout[guy == whichguy], Syout[guy == whichguy])/niter
    }
    Dn[Dn == 0] <- NA
    #cat("mean: ", mean(Dn), fill = TRUE)
    par(mar = c(3, 3, 3, 6))
    if (col == "gray") {
        cc <- seq(3, 17, , 10)/20
        cc <- rev(gray(cc))
    }
    else cc <- RColorBrewer::brewer.pal(9, "OrRd")[-1]
    image(xg, yg, Dn, col = cc)
    image.scale(Dn, col = cc)
    box()
    #return(list(grid = cbind(xg, yg), Dn = Dn))
}

gamap <- us_map(regions="counties", include="GA")
gamap <- st_transform(gamap, crs=4326) #Convert to WGS84
gamap_utm <- st_transform(gamap, crs=32617) #Convert to UTM


# Females
mat.out <- as.matrix(mc1)
M <- Mf
allS <- mat.out[,grep("sF[", colnames(mat.out), fixed=TRUE)[1:(M*2)]]
Sx <- allS[,grep(",1]", colnames(allS), fixed=TRUE)]
Sy <- allS[,grep(",2]", colnames(allS), fixed=TRUE)]
z <- mat.out[,grep("zF[", colnames(mat.out), fixed=TRUE)[1:M]]
obj <- list(Sx=Sx, Sy=Sy, z=z)

#Input coordinates are in m, output in ind/km^2 (i.e., 1e6=1000*1000)
mySCRdensity(obj, scalein=1, scaleout=1e6, col = "terrain", 
           Xl=xlim0[1],
           Xu=xlim0[2],
           Yl=ylim0[1],
           Yu=ylim0[2],
           ncolors=10)
plot(st_geometry(gamap_utm), add=TRUE, lwd = 0.25)
addnortharrow("topright", scale = 0.4)
addscalebar(pos="topleft", htin=0.07, plotepsg =32617, 
            widthhint = 0.2, label.cex = 0.6)


# Males
mat.out <- as.matrix(mc1)
M <- Mm
allS <- mat.out[,grep("sM[", colnames(mat.out), fixed=TRUE)[1:(M*2)]]
Sx <- allS[,grep(",1]", colnames(allS), fixed=TRUE)]
Sy <- allS[,grep(",2]", colnames(allS), fixed=TRUE)]
z <- mat.out[,grep("zM[", colnames(mat.out), fixed=TRUE)[1:M]]
obj <- list(Sx=Sx, Sy=Sy, z=z)

#Input coordinates are in m, output in ind/km^2 (i.e., 1e6=1000*1000)
mySCRdensity(obj, scalein=1, scaleout=1e6, col = "terrain", 
           Xl=xlim0[1],
           Xu=xlim0[2],
           Yl=ylim0[1],
           Yu=ylim0[2],
           ncolors=10)
plot(st_geometry(gamap_utm), add=TRUE, lwd = 0.25)
addnortharrow("topright", scale = 0.4)
addscalebar(pos="topleft", htin=0.07, plotepsg =32617, 
            widthhint = 0.2, label.cex = 0.6)
```

---

## Save the workspace

Save the workspace and quit
```{r saveQuit}
# Save the mcmc lists as an Rda file
save(mc1, file="mc1-bothSexes2.Rda")

# Save the entire workspace
save.image(file="Hooker_bothSexes2.Rws")

```

